---
title: "(Monte Carlo) Likelihood approximation for finite population theory"
author: "Markus Stein"
date: "July 11, 2019"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Real life meaning of probability
What are the set of all possible outcomes?  

- K. Pearson - probability distributions are observable by collecting **lots of data**. 

- W. Gosset - attempted to describe experimental design.

- R. A. Fisher - set of all possible random assignments, **permutation**.
    * 
    
### In **finite population** the meaning of probability is well estabilished.

## Finite population likelihood




## Interactive Plot

```{r eruptions, echo=FALSE}
library(shiny)
library(ggplot2)
library(plotly)



## function
allcomb <- function(seqs){    ### check efficiency using expand.grid - each step i expand.grid(1:ncurrent, vari) > then cbind()!!!!!!!!!!!!!!!!!!!!!!!!! to use in th prodALLcols function!!!
  nseq <- unlist(lapply(seqs, function(x) {
    if(is.null(dim(x))) length(x)
    else nrow(x)
  })) # number of values for each seq 
  if(is.null(dim(seqs[[1]])) && is.null(dim(seqs[[2]]))){
    row2 <- rep.int(seqs[[2]], rep.int(nseq[1],nseq[2]))
    ctab <- cbind(seqs[[1]], row2)   # current table
  } else if(is.null(dim(seqs[[1]])) && !is.null(dim(seqs[[2]]))){
    stop("under construction...")
  } else if(is.null(dim(seqs[[1]])) && is.null(dim(seqs[[2]]))){
    stop("under construction...")
  } else{
    row2 <- rep.int(1:nseq[2], rep.int(nseq[1],nseq[2]))
    ctab <- cbind(seqs[[1]][rep(1:nseq[1],nseq[2]),], seqs[[2]][row2,])  
  }
  
  if(length(seqs) == 2) {
    return( ctab ) 
  } else{
    stop("'allcomb' to 'length(seqs) > 2' under construction...")
    #     #     for(i in 3:(length(seq)-1)){
    #     for(i in 3:length(seqs)){
    #       crow = rep.int(seqs[[i]], rep.int(nrow(ctab),nseq[i]))
    #       ctab = cbind(ctab[rep(1:nrow(ctab), nseq[i]),], crow)
    #     }
  }
}

## Define UI for application that draws a histogram
ui <- fluidPage(
   
   # Application title
   titlePanel("Binary Variable Likelihood Function for Finite Population"),
   
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput("N",
                     "Population size (N):",
                     min = 1,
                     max = 30000,
                     value = 15000),
         # sliderInput("N1",
         #             "Population number of cases (N1):",
         #             min = 1,
         #             max = 30000,
         #             value = 750),
         sliderInput("p",
                     "Population probability of success (p):",
                     min = 0,
                     max = 1,
                     value = 0.5),
         sliderInput("n",
                     "Sample size (n):",
                     min = 1,
                     max = 1500,
                     value = 50),
         sliderInput("nsimMC",
                     "Monte Carlo simulation size (MC):",
                     min = 0,
                     max = 1000000,
                     value = 1000),
         actionButton("simulate", "Simulate!")
      ),
      
      # Show a plot of the generated distribution
      mainPanel(
         plotlyOutput("distPlot"),
         verbatimTextOutput("text")
      )
   )
)

## Define server logic required to draw the likelihood functions
server <- function(input, output) {
   
  observeEvent(input$simulate, {
    # generate bins based on input$bins from ui.R
    y <- rbinom(input$N,1,input$p)  # finite pop. sampled sample from Y ~ Bernoulli(p)
    U <- 1:input$N            # set of indexes for population
    S <- sample(U, input$n)   # sample of size n from U
    rangep <- seq(0,1,0.001)
    
    # dpfp <- data.frame(p = rangep, lu0=dbinom(0, 1, rangep, log=T), 
    #                    lu1=dbinom(1, 1, rangep, log=T), 
    #                    lypop=dbinom(sum(y), input$N, rangep, log=T), 
    #                    lysub=dbinom(sum(y[S]), input$n, rangep, log=T))
    lu0 <- dbinom(0, 1, rangep, log=T)
    lu1 <- dbinom(1, 1, rangep, log=T)
    # pop loglikelihood
    lypop <- dbinom(sum(y), input$N, rangep, log=T) 
    freqN <- table(y)/length(y)
    # sample loglikelihood
    lysub <- dbinom(sum(y[S]), input$n, rangep, log=T)
    freqn <- table(y[S])/length(y[S])
    # Monte Carlo loglikelihood
    MCsamp <- rbinom(input$nsimMC, input$N, freqn[2])
    ii <- allcomb(list(1:input$nsimMC, 1:length(rangep)))
    MCl <- dbinom(MCsamp[ii[,1]], input$N, rangep[ii[,2]], log=T)
    lyMC <- log(colMeans(matrix(exp(MCl), nrow = input$nsimMC, 
                                ncol = length(rangep))))
    
    dpfp <- data.frame(p = rangep, 
                       lu0 = -lu0/sum(lu0[is.finite(lu0)]), 
                       lu1 = -lu1/sum(lu1[is.finite(lu1)]), 
                       lypop = -lypop/sum(lypop[is.finite(lypop)]), 
                       lysub = -lysub/sum(lysub[is.finite(lysub)]),
                       lyMC = -lyMC/sum(lyMC[is.finite(lyMC)]))
    
    # x    <- faithful[, 2] 
    # bins <- seq(min(x), max(x), length.out = input$bins + 1)
    
    output$distPlot <- renderPlotly({

     # print(
        ggplotly(
        ggplot(dpfp, aes(rangep, lypop)) + labs(y="Standardised likelihoods") + 
      geom_ribbon(aes(ymin =  lu0 - freqn[1]/1000, ymax =  lu0 + freqn[1]/1000), fill="grey") + 
      geom_ribbon(aes(ymin = lu1 - freqn[2]/1000, ymax = lu1 + freqn[2]/1000), fill="grey") + 
      geom_line(aes(rangep, lypop, colour="Pop"), size=1) + 
      geom_line(aes(rangep, lysub, colour="Samp"), size=1) + 
      geom_line(aes(rangep, lyMC, colour="MC"), size=1) + 
      scale_colour_manual(name="LogLik" , values=c(Pop="blue", Samp="green", MC="orange")) +
      geom_vline(xintercept = input$p, color="red") + 
      geom_vline(xintercept = freqN[2], colour="blue") + 
      geom_vline(xintercept = freqn[2], colour="green")  + 
      geom_vline(xintercept = rangep[lyMC==max(lyMC)], colour="orange") #+ 
#      theme(legend.position = "bottom")
        )
      #)   
    })
    
    output$text <- renderText({
      bias <- c(freqN[2], freqn[2], rangep[lyMC==max(lyMC)]) - input$p 
      names(bias) <- c("lpop", "lsample", "lMC")
      c(paste("Bias: \n"), paste(c("lpop = ", "  ---  lsample = ", "  ---  lMC = "), format(bias, digits=2), sep=""))
    })
    
  })
  
}

## Run the application 
shinyApp(ui = ui, server = server)
```

## Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## R Output

```{r cars}
summary(cars)
```


